---
title: "Establishment probability relative time"
author: "Michael Koontz"
date: "2/26/2018"
output: 
  html_document: 
    keep_md: yes
  github_document: default
---

# Purpose

A quick outline of the establishment probability analysis when using a relative amount of time since the final introduction event.

```{r setup}
library(car)
library(ggplot2)
library(lme4)
library(lsmeans)
library(multcompView)
library(tidyr)
library(dplyr)
library(here)

b <- read.csv(here::here("/data/clean-establishment-data.csv"))
b$number <- as.factor(b$number)
b$block <- as.factor(b$block)
b$gap <- as.factor(b$gap)
```

Don't bother assessing the effect of the introduction gap, just filter out those populations

```{r filter_gap}
b.trim <-
  b %>% 
  filter(gap == FALSE)

glimpse(b.trim)
```

# Exploratory data analysis plot

Visualize how many populations were extant 3 generations after their final introduction event would get us up to Generation 7 (as we reported in the paper). Looking 5 generations is as far as we can look, given the limits of our data collection.

```{r eda_visualization_3_after}
ggplot(b.trim, aes(x = extant_3_after)) +
  geom_histogram(stat = "count") +
  facet_grid(environment ~ number)
```

```{r eda_visualization_5_after}
ggplot(b.trim, aes(x = extant_5_after)) +
  geom_histogram(stat = "count") +
  facet_grid(environment ~ number)
```


# Get simulation results

```{r get_sims}
#### Read data from the simulations ####
sims_results <- read.csv(here::here("data/simulations/simulation_stats_tidy.csv"), stringsAsFactors = FALSE)
```

# Assessment of establishment at relative time point

We will use a simple random effects structure with temporal block as a random intercept effect

## Five generations after the final introduction event. 

This uses data all the way up through generation 9 (for the 4x5 intro regime, generation 9 represents 5 generations after the final introduction event)

```{r model_5after}
prob_establishment_model <- glmer(extant_5_after ~ number*environment + (1 | block), data=b.trim, family=binomial, control=glmerControl(optimizer="bobyqa"))

Anova(prob_establishment_model)

m_5after_2 <- update(prob_establishment_model, formula= .~. - number:environment)

anova(prob_establishment_model, m_5after_2)
```

We fail to reject the null hypothesis that the model with the `number*environment` interaction equally as likely than the model without it, so we proceed by simplifying the model and dropping the interaction term.

Another way to look at group-level effects like we did with likelihood ratio tests is using a Type II Wald $\chi^2$ test. I think this test is somewhat anti-conservative, but it gives us a good ballpark and corroborates the results from our Likelihood Ratio Test.

Get simulation results averaged across environment treatments:

```{r get_sims_5after_averaged}
sims_establish_relative <- sims_results %>%
  mutate(propagule_number = substr(intro_regime, start = nchar(intro_regime), stop = nchar(intro_regime))) %>%
  filter(response == "extant_prop" & time_type == "relative") %>%
  select(propagule_number, env, t_equals_5) %>% # Here is where we average establishment across environment treatment
  rename(establish_prop = t_equals_5) %>% 
  group_by(propagule_number) %>% 
  summarize(establish_prop = mean(establish_prop)) %>% 
  as.data.frame()

```

```{r visualize_5after}
establishment_results <- lsmeans::lsmeans(prob_establishment_model, pairwise ~ number, adjust="none")
establishment_posthoc <- summary(establishment_results$lsmeans)

# Get compact letter display for significant differences across levels
establishment_sig_letters <- lsmeans::cld(establishment_results, Letters = letters, sort = FALSE, adjust = "none")$.group
# Remove white space from significance letters object
establishment_sig_letters <- gsub(establishment_sig_letters, pattern = " ", replacement = "")

establishment_xvals <- 1:4

# Minimum y value of the plot is the lowest possible value that can be plotted (lowest lower CI or 
# lowest result from simulations)
xlim <- range(establishment_xvals) + c(-0.5, 0.5)
min_y <- min(c(plogis(establishment_posthoc$asymp.LCL)), sims_establish_relative$establish_prop)

plot(x = establishment_xvals, y = plogis(establishment_posthoc$lsmean), 
     ylim=c(min_y - 0.1, 1.05), 
     xlim = xlim, 
     las=1, 
     pch = 19,
     xlab = "Introduction regime",
     ylab = "Establishment probability",
     xaxt = "n",
     main = "5 generations after final intro")

axis(side = 1, 
     at = 1:4, 
     labels = c("20x1","10x2","5x4","4x5"), 
     tick = FALSE,
     line = -0.5)

segments(x0 = establishment_xvals, 
         y0 = plogis(establishment_posthoc$asymp.LCL), 
         y1 = plogis(establishment_posthoc$asymp.UCL), 
         lwd = 2)

text(x = establishment_xvals, y = 1.07, labels = establishment_sig_letters, pos = 1)

sims_xvals <- establishment_xvals
#### Add simulation results to the plot ####
points(x = establishment_xvals, y = sims_establish_relative$establish_prop,
       pch = 17)

legend("bottomright",
       inset = c(0, -0.025),
       legend = c("microcosm", "simulation"),
       pch = c(19, 17),
       bty = "n")

```

# Three generations after final introduction event

```{r model_3after}
prob_establishment_model <- glmer(extant_3_after ~ number*environment + (1 | block), data=b.trim, family=binomial, control=glmerControl(optimizer="bobyqa"))

Anova(prob_establishment_model)

```

### Effect of interaction bw environment/intro regime
```{r interaction_3after}
m_3after_2 <- update(prob_establishment_model, . ~ . - number:environment)

anova(prob_establishment_model, m_3after_2)
```

It looks like there is a significant interaction between environment and introduction regime now. We won't be averaging our contrasts across the environment treatment this time.

### Main effect of propagule number
```{r main_effect_prop_number_3after}
m_3after_3 <- update(m_3after_2, . ~ . - number)

anova(m_3after_2, m_3after_3)

```

### Main effect of environment

```{r main_effect_env_3after}
m_3after_4 <- update(m_3after_2, . ~ . - environment)

anova(m_3after_2, m_3after_4)

```

```{r get_sims_3after_not_averaged}
sims_establish_relative <- sims_results %>%
  mutate(propagule_number = substr(intro_regime, start = nchar(intro_regime), stop = nchar(intro_regime))) %>%
  filter(response == "extant_prop" & time_type == "relative") %>%
  select(propagule_number, env, t_equals_3) %>% # Here is where we average establishment across environment treatment
  rename(establish_prop = t_equals_3) %>% 
  arrange(propagule_number, env) %>% 
  as.data.frame()

```

```{r visual_3after}
establishment_results <- lsmeans::lsmeans(prob_establishment_model, pairwise ~ environment*number, adjust="none")
establishment_posthoc <- summary(establishment_results$lsmeans)

# Get compact letter display for significant differences across levels
establishment_sig_letters <- lsmeans::cld(establishment_results, Letters = letters, sort = FALSE, adjust = "none")$.group
# Remove white space from significance letters object
establishment_sig_letters <- gsub(establishment_sig_letters, pattern = " ", replacement = "")

offset_xvals <- function(x, offset) {
  c(sapply(x, FUN = function(j) c(j - offset, j + offset)))
}

establishment_xvals <- offset_xvals(1:4, 0.2)

# Minimum y value of the plot is the lowest possible value that can be plotted (lowest lower CI or 
# lowest result from simulations)
xlim <- range(establishment_xvals) + c(-0.5, 0.5)

plot(x = establishment_xvals, y = plogis(establishment_posthoc$lsmean), 
     ylim=c(min_y - 0.03, 1.05), 
     xlim = xlim, 
     las=1, 
     pch=c(1, 19),
     xlab = "Introduction regime",
     ylab = "Establishment probability",
     xaxt = "n",
     main = "3 generations after final intro")

axis(side = 1, 
     at = 1:4, 
     labels = c("20x1","10x2","5x4","4x5"), 
     tick = FALSE,
     line = -0.5)

segments(x0 = establishment_xvals, 
         y0 = plogis(establishment_posthoc$asymp.LCL), 
         y1 = plogis(establishment_posthoc$asymp.UCL), 
         lwd = 2)

text(x = establishment_xvals, y = 1.07, labels = establishment_sig_letters, pos = 1)

#### Add simulation results to the plot ####

sims_xvals <- offset_xvals(establishment_xvals, offset = 0.05)

#### Add simulation results to the plot ####
points(x = establishment_xvals[c(1, 3, 5, 7)], y = sims_establish_relative$establish_prop[c(1, 3, 5, 7)],
       pch = 2)

points(x = establishment_xvals[c(2, 4, 6, 8)], y = sims_establish_relative$establish_prop[c(2, 4, 6, 8)],
       pch = 17)

legend("bottomleft",
       inset = c(0, -0.025),
       legend = c("fluctuating", "stable"), 
       pch = c(0, 15),
       bty = "n")

legend("bottomright",
       inset = c(0, -0.025),
       legend = c("microcosm", "simulation"),
       pch = c(19, 17),
       bty = "n")

```


# No interaction

```{r get_sims_3after_averaged}
sims_establish_relative <- sims_results %>%
  mutate(propagule_number = substr(intro_regime, start = nchar(intro_regime), stop = nchar(intro_regime))) %>%
  filter(response == "extant_prop" & time_type == "relative") %>%
  select(propagule_number, env, t_equals_3) %>% # Here is where we average establishment across environment treatment
  rename(establish_prop = t_equals_3) %>% 
  group_by(propagule_number) %>%
  summarize(establish_prop = mean(establish_prop)) %>% 
  as.data.frame()

```

```{r visual_3after_averaged}
establishment_results <- lsmeans::lsmeans(prob_establishment_model, pairwise ~ number, adjust="none")
establishment_posthoc <- summary(establishment_results$lsmeans)

# Get compact letter display for significant differences across levels
establishment_sig_letters <- lsmeans::cld(establishment_results, Letters = letters, sort = FALSE, adjust = "none")$.group
# Remove white space from significance letters object
establishment_sig_letters <- gsub(establishment_sig_letters, pattern = " ", replacement = "")

establishment_xvals <- 1:4

# Minimum y value of the plot is the lowest possible value that can be plotted (lowest lower CI or 
# lowest result from simulations)
xlim <- range(establishment_xvals) + c(-0.5, 0.5)

plot(x = establishment_xvals, y = plogis(establishment_posthoc$lsmean), 
     ylim=c(min_y - 0.03, 1.05), 
     xlim = xlim, 
     las = 1, 
     pch = 19,
     xlab = "Introduction regime",
     ylab = "Establishment probability",
     xaxt = "n",
     main = "3 generations after final intro")

axis(side = 1, 
     at = 1:4, 
     labels = c("20x1","10x2","5x4","4x5"), 
     tick = FALSE,
     line = -0.5)

segments(x0 = establishment_xvals, 
         y0 = plogis(establishment_posthoc$asymp.LCL), 
         y1 = plogis(establishment_posthoc$asymp.UCL), 
         lwd = 2)

text(x = establishment_xvals, y = 1.07, labels = establishment_sig_letters, pos = 1)

#### Add simulation results to the plot ####

sims_xvals <- establishment_xvals

#### Add simulation results to the plot ####
points(x = establishment_xvals, y = sims_establish_relative$establish_prop,
       pch = 17)

legend("bottomright",
       inset = c(0, -0.025),
       legend = c("microcosm", "simulation"),
       pch = c(19, 17),
       bty = "n")

```

The probability of establishment across introduction regimes when measured 3 generations after final introduction event are...

```{r}
data.frame(establishment_posthoc$number, plogis(establishment_posthoc$lsmean))
```

