---
title: "Population size relative time"
author: "Michael Koontz"
date: "2/26/2018"
output: 
  html_document: 
    keep_md: yes
  github_document: default
---

# Purpose

A quick outline of the population size analysis when using a relative amount of time since the final introduction event.

```{r setup}
library(car)
library(ggplot2)
library(lme4)
library(lsmeans)
library(multcompView)
library(tidyr)
library(dplyr)
library(here)

b <- read.csv(here::here("/data/clean-establishment-data.csv"))
b$number <- as.factor(b$number)
b$block <- as.factor(b$block)
b$gap <- as.factor(b$gap)
```

Don't bother assessing the effect of the introduction gap, just filter out those populations

```{r filter_gap}
b.trim <-
  b %>% 
  filter(gap == FALSE)

glimpse(b.trim)
```

# Exploratory data analysis plot

Visualize how many populations were extant 2 generations after their final introduction event would get us up to Generation 7 (as we reported in the paper). Looking 5 generations is as far as we can look, given the limits of our data collection.

```{r eda_visualization_3_after}
ggplot(b.trim, aes(x = N_3_after)) +
  geom_histogram(stat = "count") +
  facet_grid(environment ~ number)
```

```{r eda_visualization_5_after}
ggplot(b.trim, aes(x = N_5_after)) +
  geom_histogram(stat = "count") +
  facet_grid(environment ~ number)
```


# Build some models

## Five generations after final introduction event

We will use a simple random effects structure with temporal block as a random intercept effect

### Influence of fixed effects

```{r fixed_effects_interaction}
# Use LRT tests to guide interpretation, but all fixed effects will remain in the model in the end
popSize_model <- glmer(N_5_after ~ number*environment + (1 | block), data=b.trim, family=poisson, control=glmerControl(optimizer="bobyqa"))

m7 <- update(popSize_model, formula= .~. - number:environment)

anova(popSize_model, m7)
```

We reject the null hypothesis that the model with the `number*environment` interaction equally as likely than the model without it, so we keep the interaction term and will base all interpretations without averaging across any treatments.

Another way to look at group-level effects like we did with likelihood ratio tests is using a Type II Wald $\chi^2$ test. I think this test is somewhat anti-conservative, but it gives us a good ballpark and corroborates the results from our Likelihood Ratio Test.

```{r wald}
Anova(popSize_model)
```

### Interpretation and contrasts


```{r get_sims_and_samps_5after}
#### Read and prepare data from the simulations ####
sims_results <- read.csv(here::here("data/simulations/simulation_stats_tidy.csv"), stringsAsFactors = FALSE)

sims_popSize_relative <- sims_results %>%
  mutate(propagule_number = substr(intro_regime, start = nchar(intro_regime), stop = nchar(intro_regime))) %>%
  filter(response == "mean_N" & time_type == "relative") %>%
  select(propagule_number, env, mean_N = t_equals_5) %>% # Here is where we get the simulation results from GEN 7
  arrange(propagule_number, env)

#### Read samples data to put the equilibrium abundance value on the abundance plot ####
samps <- read.csv(here::here("data/NBBg-samples/NBBg-samples-combined.csv"))
equilibrium_popSize <- log(samps$R0) / samps$alpha

offset_xvals <- function(x, offset) {
  c(sapply(x, FUN = function(j) c(j - offset, j + offset)))
}
```

```{r interpretation_and_contrasts_model_5after}
popSize_results <- lsmeans::lsmeans(popSize_model, pairwise ~ environment + number, adjust="none")
popSize_posthoc <- summary(popSize_results$lsmeans)

popSize_sig_letters <- lsmeans::cld(popSize_results, Letters = letters, sort = FALSE, adjust = "none")$.group
popSize_sig_letters <- gsub(popSize_sig_letters, pattern = " ", replacement = "")
```

Visualize with both the simulation results and microcosm results

```{r interpretation_and_contrasts_plot_5after}
popSize_xvals <- offset_xvals(1:4, 0.1)

min_y <- min(c(exp(popSize_posthoc$asymp.LCL), sims_popSize_relative$mean_N))
max_y <- max(exp(popSize_posthoc$asymp.UCL))
xlim <- range(popSize_xvals) + c(-0.4, 0.4)

plot(x = popSize_xvals, y = exp(popSize_posthoc$lsmean),
     type = "n", # Set the plot up, but do not print lines yet
     ylim = c(-10, max_y + 5), 
     xlim = xlim, 
     las = 1, 
     pch = 1, 
     xaxt = "n",
     xlab = "Introduction regime",
     ylab = "Population size",
     main = "5 generations after final intro")

segments(x0 = popSize_xvals, 
         y0 = exp(popSize_posthoc$asymp.LCL), 
         y1 = exp(popSize_posthoc$asymp.UCL), 
         lwd = 2)

points(x = popSize_xvals[c(1, 3, 5, 7)], y = exp(popSize_posthoc$lsmean[c(1, 3, 5, 7)]),
       pch = 1)

points(x = popSize_xvals[c(2, 4, 6, 8)], y = exp(popSize_posthoc$lsmean[c(2, 4, 6, 8)]),
       pch = 19)

axis(side = 1, 
     at = 1:4, 
     labels = c("20x1", "10x2", "5x4", "4x5"), 
     tick = FALSE,
     line = -0.5)

legend("bottomleft",
       inset = c(0, -0.025),
       legend = c("fluctuating", "stable"), 
       pch = c(0, 15),
       bty = "n")

legend("bottom",
       inset = c(0, 0.15),
       legend = "equilibrium size", 
       lty = 2,
       lwd = 2,
       bty = "n")

legend("bottomright",
       inset = c(0, -0.025),
       legend = c("microcosm", "simulation"), 
       pch = c(19, 17),
       bty = "n")

text(x = popSize_xvals, y = max_y + 9, labels = popSize_sig_letters, pos = 1)

sims_xvals <- offset_xvals(popSize_xvals, offset = 0.05)

#### Add simulation results to the plot ####
points(x = popSize_xvals[c(1, 3, 5, 7)], y = sims_popSize_relative$mean_N[c(1, 3, 5, 7)],
       pch = 2)

points(x = popSize_xvals[c(2, 4, 6, 8)], y = sims_popSize_relative$mean_N[c(2, 4, 6, 8)],
       pch = 17)

segments(x0 = xlim[1], x1 = xlim[2], y0 = mean(equilibrium_popSize), y1 = mean(equilibrium_popSize), lty = "dashed", lwd = 2)
```

## Three generations after final intro event

This analysis will use generation 7 as the latest possible generation (for the 4x5 intro regime, the 7th generation represents 3 generations after the final introduction.)

Get the simulation results we want.

```{r get_sims_and_samps_3after}
sims_popSize_relative <- sims_results %>%
  mutate(propagule_number = substr(intro_regime, start = nchar(intro_regime), stop = nchar(intro_regime))) %>%
  filter(response == "mean_N" & time_type == "relative") %>%
  select(propagule_number, env, mean_N = t_equals_3) %>% # Here is where we get the simulation results from GEN 7
  arrange(propagule_number, env)
```

Build the model for the microcosm data.

```{r interpretation_and_contrasts_model_3after}
popSize_model <- glmer(N_3_after ~ number*environment + (1 | block), data=b.trim, family=poisson, control=glmerControl(optimizer="bobyqa"))

Anova(popSize_model)
```

### Effect of interaction

```{r interaction_effect}
m_3after_2 <- update(popSize_model, . ~ . - number:environment)

anova(popSize_model, m_3after_2)

lsmeans::lsmeans(popSize_model, pairwise ~ environment + number, adjust="none")
```

### Effect of environment

```{r env_effect}
m_3after_3 <- update(m_3after_2, . ~ . - environment)
anova(m_3after_2, m_3after_3)

lsmeans::lsmeans(popSize_model, pairwise ~ environment, adjust="none")
```


### Effect of introduction regime

```{r intro_regime_effect}
m_3after_4 <- update(m_3after_2, . ~ . - number)
anova(m_3after_2, m_3after_4)

lsmeans::lsmeans(popSize_model, pairwise ~ number, adjust="none")
```

```{r interpretation_and_contrasts_visualize_3after}
popSize_results <- lsmeans::lsmeans(popSize_model, pairwise ~ environment + number, adjust="none")
popSize_posthoc <- summary(popSize_results$lsmeans)

popSize_sig_letters <- lsmeans::cld(popSize_results, Letters = letters, sort = FALSE, adjust = "none")$.group
popSize_sig_letters <- gsub(popSize_sig_letters, pattern = " ", replacement = "")

popSize_xvals <- offset_xvals(1:4, 0.1)

min_y <- min(c(exp(popSize_posthoc$asymp.LCL), sims_popSize_relative$mean_N))
max_y <- max(exp(popSize_posthoc$asymp.UCL))
xlim <- range(popSize_xvals) + c(-0.4, 0.4)

plot(x = popSize_xvals, y = exp(popSize_posthoc$lsmean),
     type = "n", # Set the plot up, but do not print lines yet
     ylim = c(-10, max_y + 4), 
     xlim = xlim, 
     las = 1, 
     pch = 1, 
     xaxt = "n",
     xlab = "Introduction regime",
     ylab = "Population size",
     main = "3 generations after final intro")

segments(x0 = popSize_xvals, 
         y0 = exp(popSize_posthoc$asymp.LCL), 
         y1 = exp(popSize_posthoc$asymp.UCL), 
         lwd = 2)

points(x = popSize_xvals[c(1, 3, 5, 7)], y = exp(popSize_posthoc$lsmean[c(1, 3, 5, 7)]),
       pch = 1)

points(x = popSize_xvals[c(2, 4, 6, 8)], y = exp(popSize_posthoc$lsmean[c(2, 4, 6, 8)]),
       pch = 19)

axis(side = 1, 
     at = 1:4, 
     labels = c("20x1", "10x2", "5x4", "4x5"), 
     tick = FALSE,
     line = -0.5)

legend("bottomleft",
       inset = c(0, -0.025),
       legend = c("fluctuating", "stable"), 
       pch = c(0, 15),
       bty = "n")

legend("bottom",
       inset = c(0, 0.15),
       legend = "equilibrium size", 
       lty = 2,
       lwd = 2,
       bty = "n")

legend("bottomright",
       inset = c(0, -0.025),
       legend = c("microcosm", "simulation"), 
       pch = c(19, 17),
       bty = "n")

text(x = popSize_xvals, y = max_y + 9, labels = popSize_sig_letters, pos = 1)

sims_xvals <- offset_xvals(popSize_xvals, offset = 0.05)

#### Add simulation results to the plot ####
points(x = popSize_xvals[c(1, 3, 5, 7)], y = sims_popSize_relative$mean_N[c(1, 3, 5, 7)],
       pch = 2)

points(x = popSize_xvals[c(2, 4, 6, 8)], y = sims_popSize_relative$mean_N[c(2, 4, 6, 8)],
       pch = 17)

segments(x0 = xlim[1], x1 = xlim[2], y0 = mean(equilibrium_popSize), y1 = mean(equilibrium_popSize), lty = "dashed", lwd = 2)
```
